---
title: "Exercise_04_Song"
author: Yiluan Song
format: html
project:
  execute-dir: project
---

```{r setup}
library(tidyverse)
```

## Task 1: Create & Clone Repository 

I created a separate repo and added it here as a submodule.

```
git submodule add -f https://github.com/pitt-songlab/EF_Activities_Ex4.git solutions/Exercise_04/Exercise04Song
```

I also made this folder an R package.
```{r, eval=FALSE}
usethis::create_package("Exercise04Song/")
```
At this point, the working directory will automatically be changed to `Exercise04Song/`.

## Task 2: Add the first function: download.phenocam

I created a folder called `R` and added the R script there, aligning with the structure of an R package.

I named the script as `01_download_phenocam.R`, avoiding `.` in file names. Similarly, I replaced `.` with `_` in function names.

I also added `##' @export` to make sure that functions will be exported to the package.

## Task 3: Collaborator adds plot.phenocam

## Task 4: Owner adds pred.logistic and fit.logistic

Now that all R scripts are in place, I used the `roxygen2` package to write documentation and used the `devtools` package to install the package.
```{r, eval = FALSE}
roxygen2::roxygenise()
devtools::install()
```

## Task 5: Collaborator adds the master script
If this is a close collaboration, I recommend using branches instead of forks. If we are on a fork and want to pull from the upstream repo, make sure we check the name of the main branch. It might not be `master` in the command below.
```
git pull URL master
```

I added the workflow in the `vignettes` folder, aligning with the structure of an R package. The vignette was started with the following.
```{r, eval = FALSE}
usethis::use_vignette("workflow")
```

The vignette could be rendered and viewed with the follwing.
```{r, eval = FALSE}
devtools::install(build_vignettes = TRUE)
vignette(package = "Exercise04Song")
vignette("workflow", package = "Exercise04Song")
```

